### 2014-04-02 docs/paper/paper.tex
backend description and 2D oscillator

\begin{itemize}
\item VC707 board
\item resources available -- BRAM, DSP

@@

\item implemented design -- number of population units, DV buffers

\item FPGA resources utilized; maximum theoretical design size
\item clock speed and implications of this

@@

\item Ethernet communication for control and I/O -- maximum channel capacity in terms of DVs

@@

\end{itemize}

@@

@@

@@

% FIXME discussion of result?

\subsection{elaborate simulations}

>>>>

@@

% FIXME implemented design size

@@

@@

% FIXME maximum channel capacity in terms of decoded values

@@

% FIXME quick description of Nengo
Nengo's ensembles of neurons become populations that are simulated in the hardware on population units.
Connections between populations are handled through encoders and decoders.
Nodes and probes, which collect data or evaluate arbitrary functions in software, are
simulated on the host computer and are transferred to and from the board over hardware 
I/O channels.

A backend to Nengo performs the task of compiling a Nengo model to a loadfile that can be
simulated on the hardware. The compiler calculates principal components for each population
of neurons, and then performs a clustering operation in order to associate many
populations with each set of principal components while minimizing the differences
between the approximate principal components and the actual values.
The compiler calculates decoders and encoders for each population with respect
to the shared principal components, and computes an optimized encoder schedule
that coordinates reads made by each encoder so that all operations can be performed
in as few clock cycles as possible while avoiding a read collision on any decoded value block.
Once the compiler has finished writing out the loadfile, the Nengo simulator
transfers the configuration data to the hardware and controls input and output during the simulation.

@@

% integrator (1d)
% oscillator (2d)
% feed-forward nonlinear functions
% 2D multiplication

% scope of situations for which it works well
% full neurons vs. CPU populations vs. hardware populations
% high-frequency breakdown (square wave, sinusoids)
% qualitative comparison: show the figure, talk about the figure
% possible quantitative approach (RMSE)

% hardware-specific results: how many populations, 
% circumstances under which an overflow could occur
% (probabilistic analysis), power analysis,
% resource utilization

@@

Given a constant input value of 0.5, the output from the integrator approaches this value
over time and reaches it after 1.0 seconds, then continues to accumulate the input until
another second has elapsed. The population saturates and remains at 1.0 after this time
because it is not able to represent values with a magnitude greater than 1.0. However,
the integrator's behaviour for the first two seconds of simulation is correct and confirms
the proper operation of the one-dimensional population hardware.

The two-dimensional population unit was tested with a simple single-population van der Pol oscillator.
Figure~\ref{lst:oscillator2d} shows the code listing and the network generated in Nengo;
Figure~\ref{fig:oscillator2d} shows output from the \design{} hardware simulation.
\begin{figure}
\centering

\begin{subfigure}[b]{\textwidth}
\centering
\lstset{language=Python}
\begin{lstlisting}[frame=single]
import nengo
import nengo.helpers
model = nengo.Model('Oscillator')
neurons = nengo.Ensemble(nengo.LIF(200), dimensions=2, label='van der Pol Oscillator')
input = nengo.Node(output=nengo.helpers.piecewise({0:[1,0], 0.1:[0,0], label='Initial condition')
tau = 0.1
freq = 5.0
input2output = nengo.Connection(input, neurons)
recurrent = nengo.Connection(neurons, neurons, transform=[[1, -freq*tau],[freq*tau,1]], filter=tau)
input_probe = nengo.Probe(input, 'output')
neuron_probe = nengo.Probe(neurons, 'decoded_output', filter=0.005)
\end{lstlisting}
\caption{Source Code}
\label{lst:oscillator2d:code}
\end{subfigure}

\begin{subfigure}[b]{0.3\textwidth}
\centering
% FIXME \includegraphics[width=3.0in]{oscillator-2d-schematic.eps}
\caption{Schematic}
\label{lst:oscillator2d:schematic}
\end{subfigure}

\caption[A 2D neural oscillator in Nengo.]
{Source code (\ref{lst:oscillator2d:code}) and schematic (\ref{lst:oscillator2d:schematic})
of a 2D neural oscillator in Nengo.}
\label{lst:oscillator2d}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=6in]{oscillator-2d.eps}

\caption[Simulation of a 2D neural oscillator.]
{Hardware output from the simulation of a two-dimensional population representing a neural oscillator.}
\label{fig:oscillator2d}
\end{figure}

The oscillator requires a brief non-zero ``start-up'' input in order to begin oscillating.
After this input disappears (0.1~s), both dimensions of the decoded output can be seen to
oscillate in a sinusoidal fashion. This oscillation, once started, will proceed indefinitely.
% FIXME frequency analysis?
The oscillation as observed in the figure confirms the correct operation of the two-dimensional
population hardware.
